### 2018-03-13 paper/3-compute.tex
Add correctness definition

\noindent
In the above execution $\cmd{C1} \neq 1$, hence the \hs{else} branch is taken.
To extract the dependencies of \hs{select} we introduce the function \hs{track}:
a combination of \hs{execute} and \hs{dependencies} that takes the store~\store
as input and computes both the resulting value and the list of its dependencies:

@@

\noindent
The implementation uses the \hs{Writer} monad (\S\ref{sec-appendix-transformers}):
we feed \hs{fetch k = writer (store k, [k])} to the compute, which in addition
to fetching a value from the store, tracks the corresponding key. We can use
\hs{track} to track both applicative and monadic compute:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store key = if key == "A1" then 10 else 20
@\ghci@ track add store "B1"
Just (30,["A1","A2"])
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ track select store "B1"
Just (20,["C1","A2"])
\end{minted}
\vspace{1mm}

@@

\todo{AM}{Provide a formal definition of `correctness', which does not rely on materialising intermediate values.}

\subsection{Free build structures}\label{sec-free-build}

\begin{minted}{haskell}
data Depend k v r = Depend [k] ([v] -> r)

data Depends k v r = Depends [k] ([v] -> Depends k v r)
                   | Done r
\end{minted}

>>>>

\noindent
In the example above $\cmd{C1} \neq 1$, hence the \hs{else} branch is taken.
To extract the dependencies of \hs{select} we introduce the function \hs{track}:
a combination of \hs{execute} and \hs{dependencies} that takes the store~\store
as input and computes both the resulting value and the list of its dependencies:

@@

\noindent
The implementation uses the \hs{Writer} monad (\S\ref{sec-appendix-transformers}):
we feed \hs{fetch k = writer (store k, [k])} to the compute, which in addition
to fetching a value from the store, tracks the corresponding key. We can
\hs{track} both applicative and monadic compute:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store key = if key == "A1" then 10 else 20
@\ghci@ track add store "B1"
Just (30,["A1","A2"])
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ track select store "B1"
Just (20,["C1","A2"])
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store' key = if key == "A1" then 10 else 1
@\ghci@ track select store' "B1"
Just (10,["C1","A1"])
\end{minted}
\vspace{1mm}

\noindent
As expected, the dependencies of \hs{select} are determined by the contents of
the store.

% \subsection{Recognising an applicative compute wearing a monadic hat}

% Although dynamic dependencies are very useful, they typically represent only a
% small fraction of all dependencies. One can exploit this in a build system by
% recognising applicative computations even though they have the type
% \hs{Compute}~\hs{Monad} and treating them specially.

% Consider the following data type.

% \subsection{Free build structures}\label{sec-free-build}

% \begin{minted}{haskell}
% data Depend k v r = Depend [k] ([v] -> r)

% data Depends k v r = Depends [k] ([v] -> Depends k v r)
%                    | Done r
% \end{minted}

@@

Equipped with the \hs{Compute} abstraction, we can now introduce \hs{Build} in
more detail. Recall the informal definition of a build system
from~\S\ref{sec-background}:

\vspace{2mm}
\emph{Definition (Build system).}\hspace{1.5mm}Given a description of tasks, a
\emph{build system} executes them in an order that respects their dependencies
to produce a specified output from supplied inputs.
\vspace{2mm}

\noindent
To make the definition more precise, we express it as the following function:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Build c i k v = Compute c k v -> k -> Maybe i -> (k -> v) -> (i, k -> v)
\end{minted}
\vspace{1mm}

\noindent
The type closely matches the above definition: \hs{Compute c k v} is `a
description of tasks', a key~\hs{k} is~`a~specified output', and the store
\store holds `supplied inputs'. The additional input parameter \hs{Maybe i}
may provide auxiliary information from the previous build, as discussed
in~\S\ref{sec-background-summary}. The build system produces auxiliary
information \hs{i} for the next build as well as an updated store, where the
output \hs{k} must be up-to-date. We formally capture the correctness of the
resulting store in the following definition.

\definition[Correctness]{Let \hs{build :: Build c i k v} be a build system,
\hs{compute} describe an acyclic collection of build tasks, \hs{key} be an output
key, \hs{mi :: Maybe i} be auxiliary information, and \hs{store} be a store.
Then the \hs{result}ing store \hs{(_, result) = build compute k mi store}
is~\emph{correct} if there exists an \hs{ideal} store, such that

\begin{itemize}
    \item The \hs{ideal} store is \emph{consistent} with the \hs{compute}, i.e.
    for all possible non-input keys \hs{k}, the result of executing the
    \hs{compute} matches the value in the \hs{ideal} store:
    \[
    \hs{execute}~\hs{ideal}~\hs{k}~\hs{==}~\hs{Just}~\hs{(}\hs{ideal}~\hs{k)}.
    \]
    \item The \hs{result} and \hs{ideal} agree on the output \hs{key}, i.e.
    \hs{result}~\hs{key}~\hs{==}~\hs{ideal}~\hs{key}. In other words, we require
    that the output \hs{key} has the correct value, but do not impose any
    restrictions on intermediate keys, therefore permitting shallow cloud builds.
    \item The \hs{store}, \hs{ideal} and \hs{result} all agree on the inputs
    that belong to the transitive closure of \hs{key}'s dependencies. This means
    that (i) no inputs were corrupted during the build, and (ii)~the \hs{ideal}
    store has constraints not only on the output, but also on the inputs.
\end{itemize}
}
\label{def-correct}
